---
title: "Speaking Science"
subtitle: "W241 Final Project"
author: "Haerang Lee, Aris F, Mumin Khan"
date: "March 17th, 2020"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r, include=FALSE} 
# load packages 
library(data.table)
library(devtools)
library(dplyr)
library(foreign)
library(ggplot2)
library(gridExtra)
library(knitr)
library(lmtest)
library(Matching)
library(sandwich)
library(stargazer)
```



```{r}
read_data <- function(data_file) {
  if(missing(data_file)) {
    data_file <- 'speaking_science_data_03-24_clean.csv'
  }
  
  d <- fread(data_file, )
  
  names = c('response_id', 'duration_in_seconds', 'time_read_article', 'credibility', 'importance', 'q1_correct', 'q2_correct', 'q3_correct', 'q4_correct', 'q5_correct', 'q6_correct', 'questions_correct', 'time_answering_questions', 'donation', 'treatment')

  d <- d[, names, with = FALSE]
  d <- na.omit(d)
  
  #d$treatment  <- as.factor(d$treatment)
  #d$q1_correct <- as.factor(d$q1_correct)
  #d$q2_correct <- as.factor(d$q2_correct)
  #d$q3_correct <- as.factor(d$q3_correct)
  #d$q4_correct <- as.factor(d$q4_correct)
  #d$q5_correct <- as.factor(d$q5_correct)
  #d$q6_correct <- as.factor(d$q6_correct)
  
  return(d)
}
```

# EDA
We have changed the text of questions to simplify things:
- Statement_1: Please respond to the following statements. - I find this article credible.
- Statement_2: Click to write the question text - How important is it to address the pollution issue addressed in the article?
- Q_1: What is the cost to the shipping industry over the next decade to comply with the new rules? Correct answer is: Over $2 billion
- Q_2: According to the article, what is *currently* the primary source of smog-forming pollution in the state? Correct answer is: Heavy-duty diesel trucks.
- Q_3: Which of the following is NOT one of the major negative health outcomes of air pollution as mentioned in the article? Correct answer is: Bronchitis
- Q_4: Which of the following reasons is NOT a reason why establishing and enforcing a new shore power rules for ships will be difficult? Correct answer is: Plugging a ship into shore power requires many workers and several hours
- Q_5: Which statement best describes the shipment industry’s attitude toward state regulators' plan to reduce smoke-causing pollutants? Correct answer is: A new and strict regulation may be too costly for the industry.
- Q_6: According to the article, who currently bears the highest cost of pollution? Correct answer is: Residents who live near the port, in terms of negative health impacts.

```{r}
d_1 <- read_data()
head(d_1)
# Most dificult question
paste("Number of correct answers for question Q_1, is: ", sum(d_1$q1_correct))
paste("Number of correct answers for question Q_2, is: ", sum(d_1$q2_correct))
paste("Number of correct answers for question Q_3, is: ", sum(d_1$q3_correct))
paste("Number of correct answers for question Q_4, is: ", sum(d_1$q4_correct))
paste("Number of correct answers for question Q_5, is: ", sum(d_1$q5_correct))
paste("Number of correct answers for question Q_6, is: ", sum(d_1$q6_correct))
```
```{r}
boxplot(time_read_article~questions_correct, 
        data=d_1, 
        main="Number of correct answers vs. duration",
        ylab="Survey duration ", xlab="Number of correct answers out of 6 Qs ", 
        pch=19)
```


```{r}
d_1$questions_correct <- as.factor(d_1$questions_correct)
ggplot(d_1, aes(x = questions_correct, y = time_read_article, color = treatment )) + geom_boxplot()
```


```{r}
boxplot(donation~questions_correct, data=d_1, main="Number of correct answers vs. donation amount",
   ylab="Donation amount in $ ", xlab="Number of correct answers out of 6 Qs ", pch=19)
```


```{r, fig.width = 6, fig.height= 3}
d_1$treatment <- as.factor(d_1$treatment)
d_1$questions_correct <-as.factor(d_1$questions_correct)
ggplot(d_1, aes(x = questions_correct, y = donation, color = treatment )) + geom_boxplot()
```

```{r}
ggplot(
  d_1, aes(x = time_read_article, fill = treatment)) + 
  geom_density(color = "darkblue", fill = "lightblue") + 
  ggtitle("Density of Time Reading Article") + geom_rug() + 
  xlab("Seconds") + 
  ylab("Density")
```


```{r}
ggplot(d_1[treatment ==0], aes(x = donation, fill = treatment)) + geom_density(color = "darkblue", fill = "lightblue") + geom_vline(aes(xintercept = mean(donation)), color = "darkgreen", linetype = "dashed", size = 1) + geom_vline(aes(xintercept = median(donation)), color = "black", linetype = "dashed", size = 1) + xlab("Donation ($)") + annotate("text", x = 42, y = 0.02, label = "Mean: $31.11", color = "darkgreen") + ggtitle("Density of donation amount for Control Group") + geom_rug() + ylab("Density")
```
```{r}
ggplot(d_1[treatment == 1], aes(x = donation, fill = treatment)) + geom_density(color = "darkgreen", fill = "lightgreen") + geom_vline(aes(xintercept = mean(donation)), color = "darkgreen", linetype = "dashed", size = 1) + geom_vline(aes(xintercept = median(donation)), color = "black", linetype = "dashed", size = 1) + xlab("Donation ($)") + annotate("text", x = 37, y = 0.02, label = "Mean: $27.15", color = "darkgreen") + annotate("text", x = 2, y = 0.02, label = "Median: $10", color = "black") + ggtitle("Density of donation amount for Treatment Group") + geom_rug() + ylab("Density")

```
```{r}
d_1
```


```{r}
mean(d_1[d_1$treatment == 0, donation])
```


```{r,  fig.width = 7, fig.height= 3}
control_mean_donation <- mean(d_1[treatment == 0, donation])
treat_mean_donation <- mean(d_1[treatment == 1, donation])

ggplot(d_1, aes(x = donation, fill = treatment)) + 
  geom_density(alpha = 0.4) +
  geom_vline(aes(xintercept = control_mean_donation), color = "darkgreen", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = treat_mean_donation), color = "black", linetype = "dashed", size = 1) +
  annotate("text", x = 48, y = 0.02, label = paste("Control mean: $", round(control_mean_donation, 2)), color = "darkgreen") + 
  annotate("text", x = 11, y = 0.02, label = paste("Treatment mean: $", round(treat_mean_donation, 2)), color = "black") + 
  ggtitle("Density of donation amount for Control and Treatment") + geom_rug() + 
  ylab("Density") +
  xlab("Donation ($)")
```


```{r}
# Kolmogorvo-Smirnov test to check if the distributions of donations between control and treatment differ. Used the bootstrap version since 
# the distributions are not continuous
#ks.test(d_1$donation[d_1$treatment == 0],d_1$donation[d_1$treatment == 1])
ks.boot(d_1$donation[d_1$treatment == 0],d_1$donation[d_1$treatment == 1], nboots = 100)
```


```{r}
wilcox.test(d_1$donation[d_1$treatment == 0], d_1$donation[d_1$treatment == 1])
```


```{r}
hist(d_1$donation)
```


```{r}
s1<-ggplot(d_1, aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the article Credible")
s1

```

```{r}
s2<-ggplot(d_1, aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the issue discussed Important")
s2
```
Let's look at the breakdown of Credibility for the treatmnent and control groups.


```{r, fig.width = 6, fig.height= 3}
d_1$treatment <-as.factor(d_1$treatment)
d_1$importance <-as.factor(d_1$importance)
#ggplot(d_1, aes(x = importance, color = treatment )) + geom_boxplot()
```


```{r}
d_1_control <- d_1[treatment==0]
d_1_treatment <-d_1[treatment==1]
```


```{r, fig.width = 7, fig.height= 4}
plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Credibility - Control group")

plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
   xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Credibility - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```
```{r, fig.width = 7, fig.height= 4}
d_1 <- read_data()  # For some reasons I'm getting an error indicating an issue with the values supplied
scale_x_continuous()

plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Importance - Control group")

plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
   xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Importance - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```


```{r}
ks.boot(d_1$importance[d_1$treatment == 0],d_1$importance[d_1$treatment == 1], nboots = 1000)
```


```{r}
ks.boot(d_1$credibility[d_1$treatment == 0],d_1$credibility[d_1$treatment == 1], nboots = 100)
```


```{r, fig.width = 7, fig.height= 4}
d_1 <- read_data()  # For some reasons I'm getting an error indicating an issue with the values supplied
scale_x_discrete()

plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=questions_correct)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Control group")


plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=questions_correct)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```


```{r}
ks.boot(d_1$questions_correct[d_1$treatment == 0],d_1$questions_correct[d_1$treatment == 1], nboots = 100)
```


```{r,  fig.width = 7, fig.height= 3}
d_1$treatment <-as.factor(d_1$treatment)

ggplot(d_1, aes(x = time_read_article, fill = treatment)) + 
  geom_density(alpha = 0.4) +
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 0, time_read_article])), color = "darkgreen", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 1, time_read_article])), color = "black", linetype = "dashed", size = 1) +
  annotate("text", x = 105, y = 0.0035, label = "Control mean: \n 185.0 s", color = "darkgreen") + 
  annotate("text", x = 345, y = 0.004, label = "Treatm.mean: 195.9", color = "black") + 
  ggtitle("Density of time spent reading the article for Control and Treatment") + geom_rug() + 
  ylab("Density") +
  xlab("Time reading article (s)")
```


```{r}
mean(d_1[d_1$treatment == 0, time_read_article])
mean(d_1[d_1$treatment == 1, time_read_article])
```


```{r}
ks.boot(d_1$time_read_article[d_1$treatment == 0],d_1$time_read_article[d_1$treatment == 1], nboots = 100)
```


```{r}
# I think the ks.boot test is more appropriate but please confirm
t.test(d_1$time_read_article[d_1$treatment == 0], d_1$time_read_article[d_1$treatment == 1], alternative = c("two.sided"), mu = 0,  conf.level = 0.95)
```


```{r}
plot1<- ggplot(d_1_control, aes(x=donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0, 100)+
  ggtitle("Donation amount - Control group")

plot2<- ggplot(d_1_treatment, aes(x=donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0,100)+
  ggtitle("Donation amount - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```


```{r}
min_breaks <- c(0, 60, 120, 180, 240, 300, 360, 420, 480, 540, 600, 800)

hist(d_1[, time_read_article], breaks = min_breaks, freq = TRUE, labels = TRUE, 
     ylim = c(0, 60),
     xlab = "Time (Seconds)",
     main = "Time Reading the Article")
```

```{r}

minutes <- seq(1, 10) * 60

(count(d_1[time_read_article <= minutes[1]])/count(d_1))
(count(d_1[time_read_article <= minutes[2]])/count(d_1))
(count(d_1[time_read_article <= minutes[3]])/count(d_1))
(count(d_1[time_read_article <= minutes[4]])/count(d_1))
(count(d_1[time_read_article <= minutes[5]])/count(d_1))


```


# Simple Linear Regression 

```{r}
d <- read_data()
```


```{r}
mod <- lm(questions_correct ~ treatment, data=d)
stargazer(mod, type = "text")
```


```{r}
#Non parametric test
mod <- lm(credibility ~ treatment, data=d)
summary(mod)
```


```{r}
mod <- lm(importance ~ treatment, data=d)
stargazer(mod, type = "text")
```


```{r}
mod <- lm(time_read_article ~ treatment, data=d)
summary(mod)
```


```{r}
wilcox.test(d[treatment == 0, donation], d[treatment == 1, donation])
wilcox.test(d[treatment == 0, importance], d[treatment == 1, importance])
wilcox.test(d[treatment == 0, credibility], d[treatment == 1, credibility])
```


# Randomization Inference 

Testing the sharp null hypothesis that the treatment has no effect for anyone. 

```{r}
# Actual ATE's
ate_questions_correct <- d[, .('group_mean' = mean(questions_correct)), by=treatment][order(treatment), diff(group_mean)] 
ate_credibility       <- d[, .('group_mean' = mean(credibility)),       by=treatment][order(treatment), diff(group_mean)] 
ate_importance        <- d[, .('group_mean' = mean(importance)),        by=treatment][order(treatment), diff(group_mean)] 
ate_time_read_article <- d[, .('group_mean' = mean(time_read_article)), by=treatment][order(treatment), diff(group_mean)] 
ate_donation          <- d[, .('group_mean' = mean(donation)),          by=treatment][order(treatment), diff(group_mean)] 
```


```{r}
n <- 10000

# Initialize randomization inference 
ri_questions_correct <- rep(NA, n)
ri_credibility       <- rep(NA, n)
ri_importance        <- rep(NA, n)
ri_time_read_article <- rep(NA, n)
ri_donation          <- rep(NA, n)

for(i in 1:n){
  d_ri <- copy(d)
  d_ri$treatment <- sample(d_ri$treatment)
  
  # Is there any way to do this with a loop and col.names(ri)? I hate R...
  ri_questions_correct[i] <- d_ri[, .('group_mean' = mean(questions_correct)), by=treatment][order(treatment), diff(group_mean)] 
  ri_credibility[i]       <- d_ri[, .('group_mean' = mean(credibility)),       by=treatment][order(treatment), diff(group_mean)]
  ri_importance[i]        <- d_ri[, .('group_mean' = mean(importance)),        by=treatment][order(treatment), diff(group_mean)]   
  ri_time_read_article[i] <- d_ri[, .('group_mean' = mean(time_read_article)), by=treatment][order(treatment), diff(group_mean)]  
  ri_donation[i]          <- d_ri[, .('group_mean' = mean(donation)),          by=treatment][order(treatment), diff(group_mean)]  
  
}

ri <- data.table(
  questions_correct = ri_questions_correct,
  credibility       = ri_credibility,
  importance        = ri_importance,
  time_read_article = ri_time_read_article,
  donation          = ri_donation
)
```


```{r}
visualize_ri <- function(d, ate, main, xlab, ylab) {
  if (missing(main)) { 
    main <- "Histogram of d"
  }
  if (missing(xlab)) { 
    xlab <- "d"
  }
  if (missing(ylab)) { 
    ylab <- "Frequency"
  }
  
  
  h <- hist(d, 
       xlim = c(min(d)-0.25, max(d)+0.25),
       main = main,
       xlab = xlab,
       ylab = ylab
       )
  abline(v=ate, col='red', lwd = 1)

  p <- mean(abs(ate) <= d)
  text(ate, max(h$counts) - (max(h$counts)/10), paste("p = ", p))
}

visualize_ri(ri$questions_correct, ate_questions_correct, main = paste("RI(",length(ri$questions_correct),") of Questions Correct"), xlab = "Treatment Effect")
#visualize_ri(ri$credibility,       ate_credibility)
#visualize_ri(ri$importance,        ate_importance)
visualize_ri(ri$time_read_article, ate_time_read_article, main = paste("RI(",length(ri$time_read_article),") of Article Read Time (seconds)"), xlab = "Treatment Effect")
visualize_ri(ri$donation,          ate_donation, main = paste("RI(",length(ri$donation),") of Donation (USD)"), xlab = "Treatment Effect")
```



# Stargazer 
```{r echo=FALSE}
mod_1 <- lm(questions_correct ~ treatment, data=d)
mod_4 <- lm(time_read_article ~ treatment, data=d)
mod_5 <- lm(donation ~ treatment, data=d)

stargazer(mod_1,mod_4,mod_5, 
          report = ("vc*p"),
          title = "Comparing Treatment Effects",
          covariate.labels = c("Treatment"),
          omit = c("Constant", "Observations"),
          dep.var.labels = c("Questions Correct", "Article Read Time (seconds)", "Donation in USD"),
          type="html", out="models.html")
```


# Subgroup testing
```{r}
f <- read_data()
f <- f[time_read_article > 240, ]

plot(density(f$time_read_article))
```


```{r}
wilcox.test(f[treatment == 0, importance],  f[treatment == 1, importance])
wilcox.test(f[treatment == 0, credibility], f[treatment == 1, credibility])
```


```{r echo=FALSE}
g <- read_data()
g$long_read <- g$time_read_article > 120

mod_1 <- lm(questions_correct ~ treatment + long_read, data=g)
mod_2 <- lm(time_read_article ~ treatment + long_read, data=g)
mod_3 <- lm(donation ~ treatment +  long_read, data=g)

stargazer(mod_1,mod_2,mod_3, 
          report = ("vc*p"),
          title = "Comparing Treatment Effects",
          covariate.labels = c("Treatment"),
          omit = c("Constant", "Observations"),
          dep.var.labels = c("Questions Correct", "Article Read Time (seconds)", "Donation in USD"),
          type="html", out="subset_regression.html")
```


```{r echo=FALSE}

mod_1 <- lm(questions_correct ~ treatment, data=d)
mod_2 <- lm(questions_correct ~ treatment + long_read, data=g)

stargazer(mod_1,mod_2, 
          #report = ("vc*p"),
          title = "Comparing Treatment Effects",
          covariate.labels = c("Treatment", "Read time > 120"),
          omit = c("Constant", "Observations"),
          dep.var.labels = c("Questions Correct", "Article Read Time (seconds)", "Donation in USD"),
          type="html", out="subset_regression_side.html")
```


```{r}
mod <- lm(questions_correct ~ treatment + time_read_article, data=f)
stargazer(mod, type = "text")
```





