---
title: 'Speaking Science Main Survey'
output: 
    github_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_format = c('github_document')) 
    })
---

```{r include=FALSE}
# load packages 
library(foreign)
library(data.table)
library(knitr)
library(ggplot2)
library(gridExtra)
```



```{r}
d <- fread('speaking_science_data_03-24_clean.csv')
head(d)
```
We have changed the text of questions to simplify things:
- Statement_1: Please respond to the following statements. - I find this article credible.
- Statement_2: Click to write the question text - How important is it to address the pollution issue addressed in the article?
- Q_1: What is the cost to the shipping industry over the next decade to comply with the new rules? Correct answer is: Over $2 billion
- Q_2: According to the article, what is *currently* the primary source of smog-forming pollution in the state? Correct answer is: Heavy-duty diesel trucks.
- Q_3: Which of the following is NOT one of the major negative health outcomes of air pollution as mentioned in the article? Correct answer is: Bronchitis
- Q_4: Which of the following reasons is NOT a reason why establishing and enforcing a new shore power rules for ships will be difficult? Correct answer is: Plugging a ship into shore power requires many workers and several hours
- Q_5: Which statement best describes the shipment industry’s attitude toward state regulators' plan to reduce smoke-causing pollutants? Correct answer is: A new and strict regulation may be too costly for the industry.
- Q_6: According to the article, who currently bears the highest cost of pollution? Correct answer is: Residents who live near the port, in terms of negative health impacts.

Create a subset of the table with the key variables for this first EDA pass:

```{r}
names = c('Progress', 'Duration (in seconds)', 'Do you live in Los Angeles County?', 'Statement_1', 'Statement_2', 'Q_1', 'Q_2',  'Q_3', 'Q_4', "Q_5", 'Q_6', 'Win_donation', 'Q_TotalDuration', 'city', 'state', 'article' )
d_1 <- d[, names, with = FALSE]
d_1
```
```{r}
summary(d_1)
```
We have NA values in 4 key columns: "Statement_1", "Statement_2", and "Win_donation". Let's drop the NA lines from the data frame and check it again.
```{r}
d_1 <- na.omit(d_1)
d_1
```

We see that we dropped 6 entries. Let's replace the correct answers with 1 and the incorrect ones with 0 in the table above. We also create a treatment column with NY-article (control) with 0 and LA-article (treatment) with 1, in the article column.



```{r}
d_1$Q_1 <-ifelse(d_1$Q_1 == "Over $2 billion", 1, 0)
d_1$Q_2 <-ifelse(d_1$Q_2 == "Heavy-duty diesel trucks", 1, 0)
d_1$Q_3 <-ifelse(d_1$Q_3 == "Bronchitis", 1, 0)
d_1$Q_4 <-ifelse(d_1$Q_4 == "Plugging a ship into shore power requires many workers and several hours", 1, 0)
d_1$Q_5 <-ifelse(d_1$Q_5 == "A new and strict regulation may be too costly for the industry.", 1, 0)
d_1$Q_6 <-ifelse(d_1$Q_6 == "Residents who live near the port, in terms of negative health impacts.", 1, 0)
d_1$treatment <- ifelse(d_1$article == "Article-LA", 1, 0)
d_1
```



```{r}
# Most dificult question
cat("Number of correct answers for question Q_1, is: ", sum(d_1$Q_1), "\n")
cat("Number of correct answers for question Q_2, is: ", sum(d_1$Q_2),  "\n")
cat("Number of correct answers for question Q_3, is: ", sum(d_1$Q_3), "\n")
cat("Number of correct answers for question Q_4, is: ", sum(d_1$Q_4), "\n")
cat("Number of correct answers for question Q_5, is: ", sum(d_1$Q_5), "\n")
cat("Number of correct answers for question Q_6, is: ", sum(d_1$Q_6))

```

We can see that Questions 4, 2, and 3 are the most difficult ones is order of difficulty : "Which of the following reasons is NOT a reason why establishing and enforcing a new shore power rules for ships will be difficult? "


Let's make a column that sum all the correct answer for a user:

```{r}
d_1$num_cor_answ <- d_1$Q_1 + d_1$Q_2 + d_1$Q_3 + d_1$Q_4 + d_1$Q_5 + d_1$Q_6
d_1
```

```{r}
boxplot(Q_TotalDuration~num_cor_answ, data=d_1, main="Number of correct answers vs. duration",
   ylab="Survey duration ", xlab="Number of correct answers out of 6 Qs ", pch=19)
```

```{r}
boxplot(Win_donation~num_cor_answ, data=d_1, main="Number of correct answers vs. donation amount",
   ylab="Donation amount in $ ", xlab="Number of correct answers out of 6 Qs ", pch=19)
```


```{r}
plot(d_1$Q_TotalDuration, d_1$Win_donation, main="$Donation vs. survey duration",
   xlab="Survey duration ", ylab="$ Donation ", pch=19)
```

```{r}
hist(d_1$Win_donation)
```
```{r}
hist(d_1$Statement_1)
```
```{r}
s1<-ggplot(d_1, aes(x=Statement_1)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the article Credible")
s1

```

```{r}
s2<-ggplot(d_1, aes(x=Statement_2)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the issue discussed Important")
s2
```
Let's look at the breakdown of Credibility for the treatmnent and control groups.


```{r}
d_1_control <- d_1[treatment==0]
d_1_treatment <-d_1[treatment==1]
```

```{r}
plot1<- ggplot(d_1_control, aes(x=Statement_1)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  xlim(0, 7) +
  ylim(0, 55)+
  ggtitle("Importance - Control group")

plot2<- ggplot(d_1_treatment, aes(x=Statement_1)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   xlim(0, 7) +
  ylim(0, 55)+
  ggtitle("Importance - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```
```{r}
plot1<- ggplot(d_1_control, aes(x=Statement_2)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  xlim(0, 7) +
  ylim(0, 55)+
  ggtitle("Credibility - Control group")

plot2<- ggplot(d_1_treatment, aes(x=Statement_2)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   xlim(0, 7) +
  ylim(0, 55)+
  ggtitle("Credibility - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```
```{r}
plot1<- ggplot(d_1_control, aes(x=num_cor_answ)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Control group")

plot2<- ggplot(d_1_treatment, aes(x=num_cor_answ)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```

```{r}
plot1<- ggplot(d_1_control, aes(x=Win_donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0, 100)+
  ggtitle("Donation amount - Control group")

plot2<- ggplot(d_1_treatment, aes(x=Win_donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0,100)+
  ggtitle("Donation amount - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```

```{r}
# Average donation per group
d_1_control[ , mean(Win_donation)]
d_1_treatment[ , mean(Win_donation)]
```
```{r}
# Average number of correct answers per group
d_1_control[ , mean(num_cor_answ)]
d_1_treatment[ , mean(num_cor_answ)]
```


```{r}
model_1 = lm(Win_donation ~ Q_TotalDuration, data = d_1)
summary(model_1)
```














