---
title: 'Speaking Science Main Survey'
output: 
    github_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_format = c('github_document')) 
    })
---


```{r}
install.packages("Matching")
```

```{r include=FALSE}
# load packages 
library(foreign)
library(data.table)
library(knitr)
library(ggplot2)
library(gridExtra)
library(devtools)
library(dplyr)
library(Matching)
```



```{r}
d <- fread('speaking_science_data_03-24_clean.csv')
d <- d[, ID:= .I]   # adding ID column
d <-d[,c(38,1:37)]  # Moving ID column to start
```


```{r}
d
```

We have changed the text of questions to simplify things:
- Statement_1: Please respond to the following statements. - I find this article credible.
- Statement_2: Click to write the question text - How important is it to address the pollution issue addressed in the article?
- Q_1: What is the cost to the shipping industry over the next decade to comply with the new rules? Correct answer is: Over $2 billion
- Q_2: According to the article, what is *currently* the primary source of smog-forming pollution in the state? Correct answer is: Heavy-duty diesel trucks.
- Q_3: Which of the following is NOT one of the major negative health outcomes of air pollution as mentioned in the article? Correct answer is: Bronchitis
- Q_4: Which of the following reasons is NOT a reason why establishing and enforcing a new shore power rules for ships will be difficult? Correct answer is: Plugging a ship into shore power requires many workers and several hours
- Q_5: Which statement best describes the shipment industry’s attitude toward state regulators' plan to reduce smoke-causing pollutants? Correct answer is: A new and strict regulation may be too costly for the industry.
- Q_6: According to the article, who currently bears the highest cost of pollution? Correct answer is: Residents who live near the port, in terms of negative health impacts.

Create a subset of the table with the key variables for this first EDA pass:

```{r}
names = c('ID', 'duration_in_seconds', 'time_read_article', 'credibility', 'importance', 'q1_correct', 'q2_correct', 'q3_correct',
          'q4_correct', 'q5_correct', 'q6_correct', 'questions_correct', 'time_answering_questions', 'donation', 'treatment')

d <- d[, names, with = FALSE]
d
```
```{r}
summary(d)
```


```{r}
d[!complete.cases(d), ]
```




We have NA values in 2 in the importance column (IDs: 15 and 123). Let's drop the NA lines from the data frame and check it again.
```{r}
d <- na.omit(d)
d
```

We see that we dropped the 2 NA entries. 


```{r}
# Most dificult question
cat("Number of correct answers for question Q_1, is: ", sum(d$q1_correct), "\n")
cat("Number of correct answers for question Q_2, is: ", sum(d$q2_correct),  "\n")
cat("Number of correct answers for question Q_3, is: ", sum(d$q3_correct), "\n")
cat("Number of correct answers for question Q_4, is: ", sum(d$q4_correct), "\n")
cat("Number of correct answers for question Q_5, is: ", sum(d$q5_correct), "\n")
cat("Number of correct answers for question Q_6, is: ", sum(d$q6_correct))

```

We can see that Questions 4, 2, and 3 are the most difficult ones is order of difficulty 



```{r}
sum(d_1_control$Win_donation < 19)
```


```{r}
boxplot(duration_in_seconds~questions_correct, data=d, main="Number of correct answers vs. duration",
   ylab="Survey duration ", xlab="Number of correct answers out of 6 Qs ", pch=19)
```



```{r, fig.width = 6, fig.height= 3}
d_1 <-copy(d)

d_1$questions_correct <-as.factor(d_1$questions_correct)
ggplot(d_1, aes(x = questions_correct, y = duration_in_seconds, color = treatment )) + geom_boxplot()
```



```{r}
boxplot(donation~questions_correct, data=d_1, main="Number of correct answers vs. donation amount",
   ylab="Donation amount in $ ", xlab="Number of correct answers out of 6 Qs ", pch=19)
```

```{r, fig.width = 6, fig.height= 3}
d_1$treatment <-as.factor(d_1$treatment)
d_1$questions_correct <-as.factor(d_1$questions_correct)
ggplot(d_1, aes(x = questions_correct, y = donation, color = treatment )) + geom_boxplot()
```



```{r}
ggplot(d_1[treatment ==0], aes(x = donation, fill = treatment)) + geom_density(color = "darkblue", fill = "lightblue") + geom_vline(aes(xintercept = mean(donation)), color = "darkgreen", linetype = "dashed", size = 1) + geom_vline(aes(xintercept = median(donation)), color = "black", linetype = "dashed", size = 1) + xlab("Donation ($)") + annotate("text", x = 42, y = 0.02, label = "Mean: $31.11", color = "darkgreen") + ggtitle("Density of donation amount for Control Group") + geom_rug() + ylab("Density")

```
```{r}
ggplot(d_1[treatment == 1], aes(x = donation, fill = treatment)) + geom_density(color = "darkgreen", fill = "lightgreen") + geom_vline(aes(xintercept = mean(donation)), color = "darkgreen", linetype = "dashed", size = 1) + geom_vline(aes(xintercept = median(donation)), color = "black", linetype = "dashed", size = 1) + xlab("Donation ($)") + annotate("text", x = 37, y = 0.02, label = "Mean: $27.15", color = "darkgreen") + annotate("text", x = 2, y = 0.02, label = "Median: $10", color = "black") + ggtitle("Density of donation amount for Treatment Group") + geom_rug() + ylab("Density")

```
```{r}
d_1
```


```{r}
mean(d_1[d_1$treatment == 0, donation])
```


```{r,  fig.width = 7, fig.height= 3}
ggplot(d_1, aes(x = donation, fill = treatment)) + 
  geom_density(alpha = 0.4) +
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 0, donation])), color = "darkgreen", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 1, donation])), color = "black", linetype = "dashed", size = 1) +
  annotate("text", x = 48, y = 0.02, label = "Control mean: $31.11", color = "darkgreen") + 
  annotate("text", x = 11, y = 0.02, label = "Treatm.mean: $27.15", color = "black") + 
  ggtitle("Density of donation amount for Control and Treatment") + geom_rug() + 
  ylab("Density") +
  xlab("Donation ($)")
```
```{r}
# Kolmogorvo-Smirnov test to check if the distributions of donations between control and treatment differ. Used the bootstrap version since 
# the distributions are not continuous
#ks.test(d_1$donation[d_1$treatment == 0],d_1$donation[d_1$treatment == 1])
ks.boot(d_1$donation[d_1$treatment == 0],d_1$donation[d_1$treatment == 1], nboots = 100)
```

```{r}

```



```{r}
wilcox.test(d_1$donation[d_1$treatment == 0], d_1$donation[d_1$treatment == 1])
```




```{r}
plot(d_1$Q_TotalDuration, d_1$Win_donation, main="$Donation vs. survey duration",
   xlab="Survey duration ", ylab="$ Donation ", pch=19)
```

```{r}
hist(d_1$donation)
```

```{r}
s1<-ggplot(d_1, aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the article Credible")
s1

```

```{r}
s2<-ggplot(d_1, aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  scale_x_continuous(breaks=seq(0, 7, 1))+
  ggtitle("I find the issue discussed Important")
s2
```
Let's look at the breakdown of Credibility for the treatmnent and control groups.





```{r, fig.width = 6, fig.height= 3}
d_1$treatment <-as.factor(d_1$treatment)
d_1$importance <-as.factor(d_1$importance)
ggplot(d_1, aes(x = importance, color = treatment )) + geom_boxplot()
```





















```{r}
d_1_control <- d_1[treatment==0]
d_1_treatment <-d_1[treatment==1]
```


```{r, fig.width = 7, fig.height= 4}
plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Credibility - Control group")

plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=credibility)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
   xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Credibility - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```
```{r, fig.width = 7, fig.height= 4}
scale_x_continuous()

plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Importance - Control group")

plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=importance)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
   stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
   xlim(0, 7.5) +
  ylim(0, 55)+
  ggtitle("Importance - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```


```{r}
ks.boot(d_1$importance[d_1$treatment == 0],d_1$importance[d_1$treatment == 1], nboots = 1000)
```

```{r}
ks.boot(d_1$credibility[d_1$treatment == 0],d_1$credibility[d_1$treatment == 1], nboots = 100)
```





```{r, fig.width = 7, fig.height= 4}
d_1 <-copy(d)  # For some reasons I'm getting an error indicating an issue with the values supplied
scale_x_discrete()

plot1<- ggplot(d_1[d_1$treatment == 0], aes(x=questions_correct)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Control group")


plot2<- ggplot(d_1[d_1$treatment == 1], aes(x=questions_correct)) + 
  geom_histogram(binwidth = 1, color="black", fill="white")+
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  xlim(0, 7) +
  ylim(0, 30)+
  ggtitle("Correct answers - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```

```{r}

ks.boot(d_1$questions_correct[d_1$treatment == 0],d_1$questions_correct[d_1$treatment == 1], nboots = 100)

```


```{r,  fig.width = 7, fig.height= 3}
d_1$treatment <-as.factor(d_1$treatment)

ggplot(d_1, aes(x = time_read_article, fill = treatment)) + 
  geom_density(alpha = 0.4) +
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 0, time_read_article])), color = "darkgreen", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(d_1[d_1$treatment == 1, time_read_article])), color = "black", linetype = "dashed", size = 1) +
  annotate("text", x = 105, y = 0.0035, label = "Control mean: \n 185.0 s", color = "darkgreen") + 
  annotate("text", x = 345, y = 0.004, label = "Treatm.mean: 195.9", color = "black") + 
  ggtitle("Density of time spent reading the article for Control and Treatment") + geom_rug() + 
  ylab("Density") +
  xlab("Time reading article (s)")
```

```{r}
mean(d_1[d_1$treatment == 0, time_read_article])
mean(d_1[d_1$treatment == 1, time_read_article])
```

```{r}
ks.boot(d_1$time_read_article[d_1$treatment == 0],d_1$time_read_article[d_1$treatment == 1], nboots = 100)
```


```{r}
# I think the ks.boot test is more appropriate but please confirm
t.test(d_1$time_read_article[d_1$treatment == 0], d_1$time_read_article[d_1$treatment == 1], alternative = c("two.sided"), mu = 0,  conf.level = 0.95)
```


```{r}
plot1<- ggplot(d_1_control, aes(x=Win_donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0, 100)+
  ggtitle("Donation amount - Control group")

plot2<- ggplot(d_1_treatment, aes(x=Win_donation)) + 
  geom_histogram(bins=10, color="black", fill="white")+
  ylim(0,100)+
  ggtitle("Donation amount - Treatment group")

grid.arrange(plot1, plot2, ncol=2)
```

```{r}
d_1_treatment
```
```{r}
# Median donation per group
d_1_control[ , median(Win_donation)]
d_1_treatment[ , median(Win_donation)]
```

```{r}
# Average donation per group
d_1_control[ , mean(Win_donation)]
d_1_treatment[ , mean(Win_donation)]
```
```{r}
# Average number of correct answers per group
d_1_control[ , mean(num_cor_answ)]
d_1_treatment[ , mean(num_cor_answ)]
```
```{r}
# Average number of survey duration per group
d_1_control[ , mean(Q_TotalDuration)]
d_1_treatment[ , mean(Q_TotalDuration)]
```

```{r}
model_1 = lm(Win_donation ~ Q_TotalDuration, data = d_1)
summary(model_1)
```














